---
title: "Project2"
output:
  html_document:
    df_print: paged
---

#Peer-graded Assignment: Prediction Assignment Writeup


This report is for the prediction assigment writeup. The goal is to predict the classe using the data from http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.


#Background


Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.
 
##Preprocessing 

```{r}
setwd("C:\\Users\\daatas\\Desktop")
#load packages
library(caret)
library(Hmisc)
library(corrplot)
library(rattle)
library(rpart.plot)
library(randomForest)
library(gbm)

#load data
training=read.csv("pml-training.csv", header=TRUE)
testing=read.csv("pml-testing.csv", header=TRUE)

#remove na
dim(training)
training= training[, colSums(is.na(training)) == 0]
dim(training)

dim(testing)
testing= testing[, colSums(is.na(testing)) == 0]
dim(testing)

#remove empty columns
dim(training)
training=training[!sapply(training, function(x) any(x == ""))]
dim(training)

dim(testing)
testing=testing[!sapply(testing, function(x) any(x == ""))]
dim(testing)

#remove unimportant columns
training=training[,-c(1,2,3,4,5)]
testing=testing[,-c(1,2,3,4,5)]

#remove near zero variance variables

n <- nearZeroVar(training)
trainingn <- training[, -n]
testn  <- testing[,-n]
dim(trainingn)
dim(testn)

# correlations

c=cor(trainingn[,-c(1,54)])
#c
#rcorr(as.matrix(trainingn[,-c(1,54)])) #we can see that we have a small number of insignificant probabilities

corrplot(c, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

#create data for cross validation

set.seed(123456)
training = data.frame(trainingn)
inTrain <- createDataPartition(training$classe, p=0.70, list=F)
train <- training[inTrain, ]
cross_validation <- training[-inTrain, ]
```

##Prediction models

```{r}
#Decisional trees


modFit1=train(classe~.,method="rpart",data=train)
print(modFit1$finalModel)
#plot the tree
plot(modFit1$finalModel, uniform=TRUE, main="Classification Tree")
text(modFit1$finalModel, use.n=TRUE, all=TRUE, cex=.8)
fancyRpartPlot(modFit1$finalModel) 
#Confusion Matrix
confusionMatrix(predict(modFit1,newdata=cross_validation), cross_validation$classe) #using cross_validation data
#Values predicted for testing data
predict(modFit1,newdata=testing)


### Random forest

modFit<-randomForest(classe~., data=train,ntree=200)
varImp(modFit)# importance of variables
#Confusion Matrix
confusionMatrix(predict(modFit,newdata=cross_validation), cross_validation$classe) #using cross_validation data
#Values predicted for testing data
predict(modFit,newdata=testing)

### Boosted trees


modFit <- train(classe ~ ., method="gbm", data=train,trControl=trainControl(method = "repeatedcv", number = 5, repeats = 1),verbose=FALSE)
#Confusion Matrix
confusionMatrix(predict(modFit,newdata=cross_validation), cross_validation$classe) #using cross_validation data
#Values predicted for testing data
predict(modFit,newdata=testing)
```
 
